---
title: "Hyattsville Crime Data Analysis"
output: html_notebook
---

Analysis of Hyattsville City Police Department annual statistics, 2015–2025. See `data/notes.md` for source-document errors, revised figures and gaps before quoting any number.

## Setup

```{r setup, message=FALSE}
library(tidyverse)
library(scales)

totals <- read_csv("data/totals.csv", show_col_types = FALSE)
```

## Total crimes per year

Citywide totals (crimes against persons + crimes against property), each year's figure taken from its own report.

```{r citywide-totals}
citywide_totals <- totals |>
  filter(geography == "citywide", measure == "total") |>
  select(year, count)

citywide_totals
```

```{r totals-chart, fig.width=9, fig.height=5}
ggplot(citywide_totals, aes(x = year, y = count)) +
  geom_col(fill = "#2a78d6", width = 0.7) +
  geom_text(aes(label = comma(count)), vjust = -0.5, size = 3.5, color = "#52514e") +
  scale_x_continuous(breaks = 2015:2025) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.08))) +
  labs(
    title = "Total reported crimes in Hyattsville, 2015–2025",
    subtitle = "Crimes against persons plus crimes against property, citywide",
    x = NULL,
    y = "Total crimes",
    caption = "Source: Hyattsville City Police Department annual statistics reports"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold"),
    plot.caption = element_text(color = "#52514e", hjust = 0)
  )
```

**Notes on these figures**

- The 2024 total is 910 in the 2024 report but 911 in the 2025 report's prior-year column (theft revised from 464 to 465). The chart uses 910.
- The 2023 spike coincides with stolen vehicles jumping to 314. Whether offense definitions or the records system changed is an open question for HCPD.
